---
title: "Tweets Sentimental Analysis"
author: "Yatharth Malik"
date: "December 17, 2016"
output: html_document
---

## Synopsis
The analysis is aimed at knowing the sentimental behavior of people for US Election's candidates Donald J. Trump and Hillary Clinton by analysing the sentiments of their tweets.
In the analysis, we would be using Microsoft's Text Analytics API in R to calculate the sentiment score of tweets.
Here 1 indicates positive sentiment while 0 indicates negative sentiment.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the required libraries
```{r, echo=TRUE,warning=FALSE,message=FALSE}
library(twitteR)
library(httr)
library(ggplot2)
library(dplyr)
library(httr)
library(jsonlite)
```

## Setting up twitter API
Insert twitter credentials here
```{r,message=FALSE}
consumer_key = "#################################################"
access_token = "#################################################"
consumer_token = "#################################################"
acess_secret = "#################################################"

setup_twitter_oauth(consumer_key,consumer_token,access_token,acess_secret)
```

## Retrieving tweets
```{r, echo=TRUE}

number_of_tweets = 200         # Number of tweets to be extracted
trump = searchTwitter("Donald Trump+@realDonaldTrump",n = number_of_tweets ,lang = "en")
clinton = searchTwitter("Hillary Clinton + @HillaryClinton",n = number_of_tweets ,lang = "en")
trumpdf = twListToDF(trump)
clintondf = twListToDF(clinton)

# Extracting only tweets from data frame
trumpdf = select(trumpdf,text)
clintondf = select(clintondf,text)
```

## Removing duplicates and cleaning tweets
### Trump tweets
```{r,echo=T}
# Trump tweets
trump_unique = as.data.frame(trumpdf[!duplicated(trumpdf),],stringsAsFactors = F)
names(trump_unique) =  c("text")
trump_unique$text = gsub("(@|$|#)\\w+","",trump_unique$text)
trump_unique$text = gsub("http\\S+","",trump_unique$text)
trump_unique$text = gsub("RT","",trump_unique$text)
trump_unique$text = gsub("[[:punct:]]","",trump_unique$text)
trump_unique$text = gsub("\\d+\\w+","",trump_unique$text)
```
### Clinton tweets
```{r,echo=TRUE}
clinton_unique = as.data.frame(clintondf[!duplicated(clintondf),],stringsAsFactors = F)
names(clinton_unique) =  c("text")
clinton_unique$text = gsub("(@|$|#)\\w+","",clinton_unique$text)
clinton_unique$text = gsub("http\\S+","",clinton_unique$text)
clinton_unique$text = gsub("RT","",clinton_unique$text)
clinton_unique$text = gsub("[[:punct:]]","",clinton_unique$text)
clinton_unique$text = gsub("\\d+\\w+","",clinton_unique$text)
```

## Sentimental analysis of tweets
Replace ###s with Microsoft Cognitive Service's Text Analytics API key
### Trump

```{r,echo=TRUE}
trump_unique$language = "en"
trump_unique$id = seq.int(nrow(trump_unique))
trump_request = trump_unique[c(2,3,1)]
trump_size = nrow(trump_request)

trump_request_json = toJSON(list(documents = trump_request))
response = POST("https://westus.api.cognitive.microsoft.com/text/analytics/v2.0/sentiment",
                body = trump_request_json,
                add_headers(.headers = c("Content-Type"="application/json",
                "Ocp-Apim-Subscription-Key"="########################################")))

content_trump = content(response)
output_trump = data.frame(matrix(unlist(content_trump),nrow = trump_size,byrow = T),stringsAsFactors = F)
names(output_trump) = c("Sentiment_Score","ID")
output_trump$tweets = as.factor("Trump")
output_trump$Sentiment_Score = as.numeric(output_trump$Sentiment_Score)
```

### Clinton
```{r,echo = TRUE}
clinton_unique$language = "en"
clinton_unique$id = seq.int(nrow(clinton_unique))
clinton_request = clinton_unique[c(2,3,1)]
clinton_size = nrow(clinton_request)

clinton_request_json = toJSON(list(documents = clinton_request))
response = POST("https://westus.api.cognitive.microsoft.com/text/analytics/v2.0/sentiment",
                body = clinton_request_json,
                add_headers(.headers = c("Content-Type"="application/json",
                                    "Ocp-Apim-Subscription-Key"="#########################################")))

content_clinton = content(response)
output_clinton = data.frame(matrix(unlist(content_clinton),nrow = clinton_size,byrow = T),stringsAsFactors = F)
names(output_clinton) = c("Sentiment_Score","ID")
output_clinton$tweets = as.factor("Clinton")
output_clinton$Sentiment_Score  = as.numeric(output_clinton$Sentiment_Score)

```

## Plotting the result
```{r,echo=TRUE}
final = rbind(output_trump,output_clinton)
g = ggplot(final,aes(tweets,Sentiment_Score)) + 
    geom_boxplot(aes(colour = tweets)) + 
    geom_jitter(position = position_jitter(width = 0.3),colour = "grey ", alpha = 0.4) +
    labs(title = "Trump vs Clinton Tweets Analysis") + 
    labs(x = "Candidate", y = "Sentimental Score of tweets")
print(g)
```

## Summary of sentimental score of Trump's Tweets
```{r,echo=TRUE}
print(summary(output_trump$Sentiment_Score))
```

## Summary of sentimental score of Clinton's Tweets
```{r,echo=TRUE}
print(summary(output_clinton$Sentiment_Score))
```
